---
output:
  pdf_document: default
  html_document: default
---
---
title: "AMLHC-final-exam"
output: html_notebook
---
by Alexandra Schwaiger

Dataset: Hepatitis C (Kaggle)

Goal: Prediction of Hepatitis C based on 12 attributes of blood donors and Hep C patients laboratory and demographic values.

Method:
Since my dataset contains information if patients have Hepatitis or not, so I will use supervised learning methods.The task is a classification problem and my target attribute for classification is a Category (blood donors vs. Hepatitis C, including its progress: 'just' Hepatitis C, Fibrosis, Cirrhosis).
Possible Methods are random forest, SVM, NN or alternativly K-Nearest Neighbors.

needet library
```{r}
library(gridExtra)
library(randomForest)
library(tidyverse)
library(gplots)
library(readr)
library(caret)
library(GGally)
library(corrplot)
library(caTools)
library(keras)
library(reshape2)

```

1. Read in the data and get an overview

```{r}
hepc <- read.csv("C:/Users/PC/Desktop/HepatitisCdata.xls", header = TRUE) 
str(hepc)
```
615 observations and 14 attributes of blood donors and Hepatitis C patients laboratory and demographic values
All attributes except Category and Sex are numerical
The dataset contains a number of NA= Not available values

Visualisation:
```{r}

ggplot(hepc, aes(x = Category)) +
  geom_bar() +
  labs(x = "Category") +
  ggtitle("Count Plot for Category")
```
Dataset is unbalanced 


```{r}
hep_plot <- hepc[,2:14]
hep_plot$Category <- as.integer(factor(hep_plot$Category, levels = c('0=Blood Donor', '0s=suspect Blood Donor', '1=Hepatitis', '2=Fibrosis', '3=Cirrhosis'), labels = c(0, 0, 1, 2, 3)))
hep_plot$Sex <- as.integer(factor(hep_plot$Sex, levels = c('m', 'f'), labels = c(1, 0)))

numeric_vars <- names(hep_plot)[sapply(hep_plot, is.numeric)]
cor_matrix <- cor(hep_plot[, numeric_vars], use = "complete.obs")

cor_matrix_melted <- melt(cor_matrix)

ggplot(cor_matrix_melted, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "orange") +
  scale_fill_gradient2(low = "white", mid = "peachpuff2", high = "deeppink1", midpoint = 0,
                       limits = c(-1, 1), name = "Correlation") +
  geom_text(aes(label = round(value, 2)), vjust = 1) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Correlation Heatmap",
       x = "Variables", y = "Variables") +
  coord_fixed()

```


2. Split data into training and test sets
```{r}
set.seed(46)
trainIndex <- createDataPartition(hepc$Category, p = 0.75, list = FALSE)
trainData <- hepc[trainIndex, ]
testData <- hepc

```

Preprocessing Training data: Change the categorical Variables to levels and fill missing values with input of median.
```{r}

trainData$Category <- as.integer(factor(trainData$Category, levels = c('0=Blood Donor', '0s=suspect Blood Donor', '1=Hepatitis', '2=Fibrosis', '3=Cirrhosis'), labels = c(0, 0, 1, 2, 3)))
trainData$Sex <- as.integer(factor(trainData$Sex, levels = c('m', 'f'), labels = c(1, 0)))
trainData$Category <- as.factor(trainData$Category)

preProcessValues <- preProcess(trainData[, -which(names(trainData) == "Category")], method = c("medianImpute", "center", "scale"))
trainData[, -which(names(trainData) == "Category")] <- predict(preProcessValues, trainData[, -which(names(trainData) == "Category")])
```

Same for test data
```{r}
testData$Category <- as.integer(factor(testData$Category, levels = c('0=Blood Donor', '0s=suspect Blood Donor', '1=Hepatitis', '2=Fibrosis', '3=Cirrhosis'), labels = c(0, 0, 1, 2, 3)))
testData$Sex <- as.integer(factor(testData$Sex, levels = c('m', 'f'), labels = c(1, 0)))
testData$Category <- as.factor(testData$Category)
testData[, -which(names(testData) == "Category")] <- predict(preProcessValues, testData[, -which(names(testData) == "Category")])
```


Split the set into trainings- and test-set.
```{r}
x_train <- trainData %>% select(-Category)
y_train <- trainData$Category
x_test <- testData %>% select(-Category)
y_test <- testData$Category


```


Train random forest model

```{r}
model_rf <- train(x_train, y_train, method="rf", preProcess=c())

featVar <- apply(x_train, 2, var)
length(featVar<0.001)

trControl <- trainControl(method="cv", number=10)
model_rf <- train(x_train, y_train, method="rf", preProcess=c("center", "scale"),
trainControl=trControl)
```

```{r}
vi <- varImp(model_rf)
```

```{r}
best_rf <- model_rf$finalModel
y_predicted <- predict(best_rf, x_test)
confusionMatrix(y_predicted, y_test)
```
The model achieved an very good overall accuracy of 99.67%, with a Kappa value of 0.98.25. 
``



